---
title: "How Does Global Health Governance Look Like?"
author: "James Hollway, Henrique Sposito, Jael Tan, Bernhard Bieri, and Esther Peev"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How Does Global Health Governance Look Like?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Manyhealth Data

### Agreements

`{manyhealth}` package is composed of multiple databases, each of them representing a different aspect of global health governance. The **agreements** database is an ensemble of data on international health instruments. The combination of datasets that are contained in this database prints an overview of all international instruments that governs the global health sphere. The **Texts** database assembles treaty texts from the **Agreements** database. 

```{r data, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(tidyverse)
library(knitr)
head(manyhealth::agreements$WHO[,c(1:5)]) %>% 
  kable("html")

head(manyhealth::agreements$GHHR[,c(1:5)]) %>% 
  kable("html")

```

### Organizations

The **Organizations** database lists different actors playing a role in the global health governance, such as NGOs, IGOs or associations with their date of establishment and headquarter location.

```{r, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(tidyverse)
library(knitr)
head(manyhealth::organizations$CHATHAM) %>% 
  kable("html")

```


## Visualizing Global Health Governance

This graph below represents a sample of treaties from the agreements database and how they are linked to each other. The dataset selected to this end, GHHR_REF, also includes agreements on human rights. For more information on how to read the manyID, please read the [vignette](https://globalgov.github.io/manypkgs/articles/agreements.html) from `{manypkgs}` package.

```{r lineage, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(dplyr)
library(ggplot2)
library(migraph)

# Get dataset
references <- manyhealth::references$GHHR_REF %>% 
  dplyr::distinct() %>%
  dplyr::mutate(year = stringr::str_extract(Treaty2, "[:digit:]{4}")) %>%
  dplyr::filter(RefType == "Cites") %>%
  dplyr::arrange(year)

# Plot with migraph
migraph::gglineage(references[75:100,c(1,2,3)]) +
  labs(title = "Treaty Lineage of Selected Agreements from GHHR Dataset",
       caption = "Source: manyhealth") +
  theme(plot.title = element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))
```

The second graph below illustrate lineages between a sample of treaties from World Health Organization (WHO) dataset. It is mostly resolutions, decisions and conventions adopted under the auspice of WHO. 

```{r lin, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(dplyr)
library(ggplot2)
library(migraph)

# Get dataset
references <- manyhealth::references$WHO_REF %>% 
  dplyr::distinct() %>%
  dplyr::mutate(year = stringr::str_extract(Treaty2, "[:digit:]{4}")) %>%
  dplyr::filter(RefType == "Cites") %>%
  dplyr::arrange(year)

# Plot with migraph
migraph::gglineage(references[10:35,c(1,2,3)]) +
  labs(title = "Treaty Lineage of Selected Agreements from WHO",
       caption = "Source: manyhealth") +
  theme(plot.title = element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))
```

