---
title: "How Does Global Health Governance Look Like?"
author: "Esther Peev"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How Does Global Health Governance Look Like?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Manyhealth Data

`{manyhealth}` package is composed of multiple databases, each of them representing a different aspect of global health governance.

### Agreements

The **agreements** database is an ensemble of data on international health instruments. The combination of datasets that are contained in this database prints an overview of all international instruments that governs the global health sphere. With the agreements data, we can for example see all the health treaties signed in a specific year, or all WHO instruments adopted during a year.

```{r agreements, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(dplyr)
library(knitr)
manyhealth::agreements$WHO[,c(1:5)] %>% 
  dplyr::filter(Beg == "2010") %>% 
  kable("html")

manyhealth::agreements$GHHR[,c(1:5)] %>% 
  dplyr::filter(Beg == "1948") %>% 
  kable("html")

```

### Texts

The **Texts** database assembles treaty texts from the **Agreements** database. Functions from `{manypkgs}` can then be used to analysis treaty texts, some specific articles or preambles. We can, for example, get the preambles from all agreements that mention the word "HIV".

```{r texts, eval = FALSE}
manypkgs::get_articles(manyhealth::texts$AGR_TXT$TreatyText, article = "preamble", match = "HIV")

```


### Organizations

The **Organizations** database lists different actors playing a role in the global health governance, such as NGOs, IGOs or associations with their date of establishment and headquarter location. For example, the table below shows the organizations that are based in London.

```{r organizations, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(dplyr)
library(knitr)
manyhealth::organizations$CHATHAM %>%
  dplyr::filter(City == "London") %>% 
  kable("html")

```


### References

The **References** database was hand coded from TEXTS and identifies which treaties cite other treaties. With this information we can, for example, get to treaty lineages. The code below illustrates how to extract sample of treaties from the references database and how we can use {migraph} to plot treaties that cite to other treaties. The dataset selected to this end, GHHR_REF, also includes agreements on human rights.

```{r lineage, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(dplyr)
library(ggplot2)
library(migraph)

# Get dataset
references <- manyhealth::references$GHHR_REF %>% 
  dplyr::distinct() %>%
  dplyr::mutate(year = stringr::str_extract(Treaty2, "[:digit:]{4}")) %>%
  dplyr::filter(RefType == "Cites") %>%
  dplyr::arrange(year)

# Plot with migraph
migraph::gglineage(references[75:100,c(1,2,3)]) +
  labs(title = "Treaty Lineage of Selected Agreements from GHHR Dataset",
       caption = "Source: manyhealth") +
  theme(plot.title = element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))
```


The second graph below illustrate lineages between a sample of treaties from World Health Organization (WHO) dataset. It is mostly resolutions, decisions and conventions adopted under the auspice of WHO. 

```{r lin, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
library(dplyr)
library(ggplot2)
library(migraph)

# Get dataset
references <- manyhealth::references$WHO_REF %>% 
  dplyr::distinct() %>%
  dplyr::mutate(year = stringr::str_extract(Treaty2, "[:digit:]{4}")) %>%
  dplyr::filter(RefType == "Cites") %>%
  dplyr::arrange(year)

# Plot with migraph
migraph::gglineage(references[10:35,c(1,2,3)]) +
  labs(title = "Treaty Lineage of Selected Agreements from WHO",
       caption = "Source: manyhealth") +
  theme(plot.title = element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))
```

For more information on how to read the manyID, please read the [vignette](https://globalgov.github.io/manypkgs/articles/agreements.html) from `{manypkgs}` package. For more information on some of our other "many" packages, please see [manydata](https://github.com/globalgov/manydata).
